---
title: Iterating with names
author: ''
date: '2018-12-19'
slug: iterating-with-names
categories:
  - R
  - Code
tags:
  - R
  - purrr
  - tidyverse
type: post
---

# The Problem

I've come across this problem a few times lately, when I've wanted to iterate through some sort of named list or vector, and use both the name and value in each iteration. 

To illustrate, here's a vector, which I have creatively named `myvec`.  It is a numeric vector containing the numbers 1 to 26, and each element has a name, which in this case is represented by a letter of the alphabet.

```{r}
library(purrr)
myvec <- 1:26
names(myvec) <- LETTERS
myvec
```

What if I want to iterate through this vector and produce a string, containing both the value and its name?

There are lots of ways to do this, but I want something that feels clean and simple, preferably using base R or tidyverse functions.

If I attempt to solve this problem using `lapply()`, I quickly realise this won't work as `lapply()` doesn't keep the names of the object being iterated over.

```{r}
lapply(myvec, function(value){
  paste(names(value), " = ", value)
})
```

This question had been asked previous [on Stack Overflow](https://stackoverflow.com/questions/9469504/access-and-preserve-list-names-in-lapply-function) and the next thing I attempted was to use `sapply()`, which didn't work either.

```{r}
sapply(myvec, function(value){
  paste(names(value), " = ", value)
}, simplify = FALSE, USE.NAMES = TRUE)
```

The solution I went with in the end uses `purrr` and gives me further incentive to get round to learning it properly future!

My fantastic colleague, John Drummond, pointed out that this can be achieved via the use of `map()` like so:

```{r}
map(names(myvec),~paste(.x,"=",myvec[[.x]]))
```

However, I prefer the syntax of `map2()` which provides me with a slightly more readable solution:

```{r}

map2(myvec, names(myvec), ~paste(.y, "=", .x))
```

Do you know any other ways of achieving my goals: iterating through a named list or vector, using both names and values, with an emphasis on readability?  Let me know [on Twitter](https://twitter.com/nic_crane).

[Update]

As ever, the R Twitter community is amazing, and in a matter of minutes this even better `purrr` solution was highlighted by the fantastic [\@rensa_co](https://twitter.com/rensa_co) (and just minutes later [\@romain_francois](https://twitter.com/romain_francois) too!)
```{r}

imap(myvec, ~paste(.y, "=", .x))
```

A little bit of context here; bear in mind that I haven't provided argument names above, simply for the sake of brevity and readability:

`imap()` is an indexed map and its first argument `.x` (in this case `myvec`) is the list or vector.

The second argument (`.f`) can be specified in a few different ways (as a function, formula, or vector), but here I've used the formula syntax (as marked by the `~`).

Within the body of the formula, `.x` refers to the element itself and `.y` refers to the name of the element.