---
title: How do I make my own dplyr-style functions?
author: Nic Crane
date: '2018-04-16'
slug: how-do-i-make-my-own-dplyr-style-functions
categories: ['R']
tags:
  - dplyr
  - tidyeval
  - tidyverse
---

```{r, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)

suppressMessages(library(dplyr))
suppressWarnings(library(rlang))
```

In this series of blog posts introducing tidy eval, we've been looking at [why tidy eval is important](https://thisisnic.github.io/2018/03/29/what-is-tidy-eval-and-why-should-i-care/), and terms like ["quotation" and "quasiquotation"](https://thisisnic.github.io/2018/03/31/what-the-heck-is-quasiquotation/).

The next step is to look at how we can write our own dplyr-style functions in R.

This post will look at the following terms and functions:

* quosures
* `quo()`
* `enquo()`


## What is a quosure?

Quosures are a topic which come up frequently when talking about tidy eval.  But what is a quosure?  From [here](https://www.rdocumentation.org/packages/rlang/versions/0.1.6/topics/quosure)

>> Quosures are quoted expressions that keep track of an environment

On it's own, this definition is a little abstract.  We can guess from its name that the term "quosure" is related to "closure", meaning a function which keeps track of its parent environment, but in the case of "quosure", we're referring specifically to quoted expressions instead of functions.

In order to give it a little more context, let's look at an example of where you might need to use a quosure.

## Using dplyr Inside Custom Functions

In a number of previous posts, we looked at how we can use `sym()` and `!!` to help us work with *dplyr* functions when called inside our own functions.  As a recap, here's an example of a function, `wrangle_data()`, which selects a column specified by the user, filters based on a user-supplied value, and then returns the first 6 lines of the resulting dataset:

```{r}
library(dplyr)
library(rlang)

wrangle_data <- function(x, col_name, val){
  
  one_col <- select(x, !!sym(col_name))
  filtered <- filter(one_col, !!sym(col_name) == val)
  head(filtered)
}

wrangle_data(iris, "Species", "versicolor")
```

This works fine, although you may notice that the syntax is a little different to if you were to use *dplyr* functions directly; for example, when we use `select()` to return specified columns, we don't need to put the column names in quotes.

```{r}
select(iris, Species) %>%
  head()
```

So, how can we write our own functions that work in the same way?

## Capturing User Input

Let's take a quick look at what happens if we try to call our function without putting quote marks around the column name.

```{r}
wrangle_data(iris, Species, "versicolor")
```

We get an error message.  This is because R is trying to find an object named `Species`, but no object with this name exists.  We need some way of capturing the expression that the user has input and evaluating it in the correct context.

### Quosures: quo() and enquo()

This brings is back to our earlier definition of quosures:

>> Quosures are quoted expressions that keep track of an environment

We might try using `quo()` to create a quosure.  Note that we have also removed the call to `sym()`.

```{r}
wrangle_data <- function(x, col_name, val){
  
  col_name <- quo(col_name)
  
  one_col <- select(x, !!col_name)
  filtered <- filter(one_col, !!col_name == val)
  head(filtered)
}

wrangle_data(iris, Species, "versicolor")

```

This is almost correct, but not quote there!  To make this function run, we need to use `enquo()`.

```{r}
wrangle_data <- function(x, col_name, val){
  
  col_name <- enquo(col_name)
  
  one_col <- select(x, !!col_name)
  filtered <- filter(one_col, !!col_name == val)
  head(filtered)
}

wrangle_data(iris, Species, "versicolor")

```

So what is the difference between `quo()` and `enquo()` then?

### Comparing quo() and enquo()

When we call `quo()` on an expression, it takes that expression and wraps it in a quosure, in this case, returning `col_name` as a quosure.  However, what we actually wanted to do was return `Species` as a quosure, and this extra step can be performed by using `enquo()` instead.

From [the programming vignette from dplyr](https://dplyr.tidyverse.org/articles/programming.html):

>> enquo() uses some dark magic to look at the argument, see what the user typed, and return that value as a quosure. 

To look at it another way, if we wanted to turn a variable we supplied ourself into a quosure, we'd probably use `quo()`, but if it's one supplied by the user, then we should use `enquo()`.

To conclude, `enquo()` is crucial if we wish to create our own *dplyr* style functions which capture user input.  Any feedback on this post is welcome [via Twitter](https://twitter.com/nic_crane/status/985767823157923842).